---
title: "HW 4"
author: "Samantha Grayson"
date: "2023-11-14"
output: html_document
---

# Load libraries

```{r}
library(cobalt)
library(kableExtra)
library(stargazer)
library(ivreg)
library(tidyverse)
```

# Import data

```{r}
crime <- read.csv("~/Documents/GitHub/Research-Methods-II/HW 4/crime-iv.csv")
```


# 1. What implicit claim about causality does Obama's "cycle of crime" theory assert? 

Obama has coined the term "cycle of crime" to describe the phenomenon that harsh sentencing creates more criminals, which reduces the benefits of convictions and criminal sentences. The implicit claim about causality that this statement makes is that the longer someone is in jail (the harsher the sentence), the more likely it is that they will go back to jail. 

# 2. Your friend's proposed research design is: Run a regression whose outcome is recidivism and whose main explanatory variable is the length of the prison sentence. React your friend's research design. 

This regression would tell us if there's a correlation between the length of the prison sentence and recidivism. It would tell us whether or not there is a significant relationship observed between length of prison stay and whether or not the defendant goes back to jail. However, the proposed model does not allow us to make a causal claim. That is what instrumental variables, with random assignment, should allow us to do. 


# 3. Develop a separate research design

In learning that defendants are randomly assigned to judge, it appears that we can implement a different research design that would allow us to make a causal claim. We can use an Instrumental variable design where: 

Instrumental Variable: Judge's political affiliation assignment
Treatment (Nudge): Length of prison stay
Outcome: Return to jail (Recidivism)


# 4. Perform a balance test. Does the judge's party really seem to be randomly assigned? 

```{r}
covs <- subset(crime, select = c(Severity.Of.Crime))

crime$p.score <- glm(Republican.Judge ~ Severity.Of.Crime,
                     data = crime,
                     family = "binomial")$fitted.values

crime$att.weights <- with(crime, Republican.Judge + (1-Republican.Judge)*p.score/(1-p.score))

bal.tab(covs, 
        treat = crime$Republican.Judge, 
        weights = crime$att.weights,
        disp = c("means", "sds"), un = TRUE, 
        stats = c("mean.diffs", "variance.ratios"),
        thresholds = c(m = .1))
```

The output of the balance table tells us that severity of crimes are randomly assigned to judge. The mean difference between severity of crime for cases with republican judges and severity of crime for cases with democrat judges is small -- the table is balanced.  


## Make publication-worthy table for the balance table output

```{r}
balance.table <- bal.tab(covs, 
        treat = crime$Republican.Judge, 
        weights = crime$att.weights,
        disp = c("means", "sds"), un = TRUE,
        stats = c("mean.diffs", "variance.ratios"),
        thresholds = c(m = .1))

table <- balance.table$Balance

table_output <- table %>% 
  select(c(M.0.Un, SD.0.Un, M.1.Un, SD.1.Un, Diff.Un, M.Threshold)) %>% 
  dplyr::rename("Democrat Judge Unadj. Mean" = "M.0.Un") %>% 
  dplyr::rename("Democrat Judge Unadj. SD" = "SD.0.Un") %>% 
  dplyr::rename("Republican Judge Unadj. Mean" = "M.1.Un") %>% 
  dplyr::rename("Republican Judge Unadj. SD" = "SD.1.Un") %>% 
  dplyr::rename("Mean Difference Unadj." = "Diff.Un") %>% 
  dplyr::rename("Mean Threshold (.1)" = "M.Threshold") 

rownames(table_output) = c("Severity of Crime")
  
  
kable(table_output,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = c(2, 2, 2, 2, 2, 2),
      caption = "Balance Table Comparing Democrat and Republican Judge Cases") %>%
kable_classic() %>% 
cat(., file = "~/Documents/GitHub/Research-Methods-II/HW 4/balance_table_output.html")

```

Cool! We have a table output!


# 5. Describe in words the "first stage" of the IV design. Then, create a publication-quality table for the first stage only. 

Stage 1 of the Instrumental Variables design is regressing compliance on nudge. For this design, compliance is time in jail and nudge is whether the judge is democrat or republican. I will also control for severity of crime in the model.

```{r}
model1 <- lm(Months.In.Jail ~ Republican.Judge + Severity.Of.Crime,
             data = crime)

summary(model1)
```


## Make publication-worthy table 

```{r}
model_results1 <- stargazer(model1, 
                            title="First Stage", 
                            type = "html", 
                            dep.var.labels=c("Months in Jail"), 
                            covariate.labels=c("Republican Judge (1)",
                                               "Severity of Crime")) 

capture.output(model_results1, file = "~/Documents/GitHub/Research-Methods-II/HW 4/stage1.html")
```


# 6. Interpret the coeffocoent on your instrument from the first stage

Net of severity of the crime, when someone is assigned to a Republican judge, then they will spend 3.2219 more months in jail than someone assigned to a Democrat judge, on average.


# 7. Calculate the "reduced form." 

```{r}
model2 <- lm(Recidivates ~ Republican.Judge + Severity.Of.Crime,
             data = crime)

summary(model2)
```


## Make publication-worthy table 

```{r}
model_results2 <- stargazer(model2, 
                            title="Reduced Form", 
                            type = "html", 
                            dep.var.labels=c("Recidivates"), 
                            covariate.labels=c("Republican Judge (1)",
                                               "Severity of Crime")) 

capture.output(model_results2, file = "~/Documents/GitHub/Research-Methods-II/HW 4/reducedform.html")
```


# 8. Calculate the ratio of the reduced form

```{r}
0.142664 / 3.2219
```


# 9. Now complete the IV regression and make a publication quality table of the second stage. Use the setup below. 

```{r}
iv1 = ivreg(Recidivates ~ Months.In.Jail + Severity.Of.Crime | Republican.Judge + Severity.Of.Crime, data = crime)

summary(iv1)
```

## Make publication-worthy table 

```{r}
model_results3 <- stargazer(iv1, 
                            title="IV Regression", 
                            type = "html", 
                            dep.var.labels=c("Recidivates"), 
                            covariate.labels=c("Months in Jail",
                                               "Severity of Crime")) 

capture.output(model_results3, file = "~/Documents/GitHub/Research-Methods-II/HW 4/IV1.html")
```


# 10. State the F-stat in your writeup. It does not need to go into your table (although, in an actual publication it would). Is it above the conventional threshold? 

The Wald test-statistic from the model (R does not output an F-statistic) is 164.3. The statistic is significant (p < .001), and above the threshold we discussed in class of 10. 

I also used Stata to derive the F-statistic (See FStat-Calc.do in the repo). According to the output, F(2, 4997) = 164.34, confirming the output from the model ran in R. The code also outputs a weak identification test (Cragg-Donald Wald F statistic), which is 76.868. 

Both values are above the conventional threshold of 10. 


# 11. Compare your answer to question #8 (above) to the IV coefficient in #9. 

The coefficients are the same. Both are 0.04428. 

# 12. Complete these sentences.

12a. In the research design above (using randomized judges), the always-takers are the defendants who are always given long prison sentences no matter if they are assigned a democrat or republican judge, holding severity of the crime constant. 


12b. The never-takers are the defendants who are always given short prison sentences no matter if they are assigned a democrat or republican judge, holding severity of the crime constant. 


12c. The compliers are the defendants who are given long prison sentences only if they are assigned to a republican judge, holding severity of the crime constant.


12d. The defiers are the defendants who are given short prison sentences only if they are assigned to a republican judge, holding severity of the crime constant. 


# 13. Comment on the monotonicity assumption and the possibility of "defiers" in this setting. 

The monotonicity assumption says that there should not be defiers in the data. To use instrumental variables, the monotonicity assumption should be met. However, I would argue that there is a possibility of defiers in this setting. It is possible that if a defendant were to be assigned to a democrat judge, they would have gotten a longer sentence (or the opposite: if a defendant were to be assigned to a republican judge, they would have gotten a shorter sentence). 


# 14. In your dataset, what types of defendants are compliers? 

The compliers are the defendants who were given long sentences when assigned to a republican judge. 


# 15. Does the cycle of crime hypothesis appear to be true for the compliers? 

Yes - defendants who are given a longer sentence (when assigned to a republican) are more likely to recidivate (beta coefficient = 0.044, p < .001), controlling for the severity of the crime committed.







