---
title: "HW 2 Code"
author: "Samantha Grayson"
date: "2023-10-31"
output: html_document
---

# 1. Load the data

```{r}
# Load packages
library(stargazer)
library(survey)
library(tidyverse)

# Load data
data <- read.csv("~/Documents/GitHub/Research-Methods-II/HW 2/vaping-ban-panel.csv")
```

# 2. Use a regression to evaluate the "parallel trends" requirement of a difference-in-difference ("DnD") estimate.


```{r}
# Data clean-up

data <- data %>% 
  group_by(State.Id) %>% 
  mutate(Vaping.Ban = as.numeric(Vaping.Ban)) %>% 
  mutate(temp = sum(Vaping.Ban)) %>% 
  mutate(treatment = ifelse(temp >0, 1, 0)) %>% 
  ungroup() %>% 
  select(-temp) %>% 
  mutate(treatment = as.factor(treatment))%>% 
  mutate(State.Id = as.factor(State.Id))

# Make a pre/post variable

data <- data %>% 
  mutate(post = ifelse(Year < 2021, 0, 1)) %>% 
  mutate(post = as.factor(post))

## NOTE: If post = 0, that means it is pre-treatment

```


```{r}
# Limit dataset to all pre-treatment observations
data_regression1 <- data %>% 
  filter(post == 0)

# Compare pres of treatment versus non-treatments
model1 <- lm(Lung.Hospitalizations ~ treatment + Year + treatment * Year, data = data_regression1)

summary(model1)

# Is the interaction significant? 

```
The interaction between treatment and year pre-2021 is not significant, so we can say that the parallel trend requirement has been satisfied. 


# 3. Create the canonical DnD line graph. 

```{r}

ggplot(data = data,
       mapping = aes(x = Year,
                     y = Lung.Hospitalizations,
                     color = treatment)) +
  geom_line()+
  geom_vline(xintercept = 2021, linetype="dotted", 
                color = "red", size=1.5)+
  ylab("Lung Hospitalizations")  + 
  scale_color_discrete(name = "Vaping Ban Implemented", labels = c("No", "Yes"))

```

This graph shows that pre-treatment, the lines in non-treated (No Vaping Ban) and treated (Vaping Ban) places are similar (parallel line assumption is met). However, after the onset of treatment in 2021, the lines for treated and non-treated places change course. The line for treated places drops, demonstrating that the ban seems to decrease lung hospitalizations. Further testing is required to see if the results are significant. 


# 4. Run a regression to estimate the treatment effect of the laws

```{r}
model2 <- lm(Lung.Hospitalizations ~ post + treatment + post * treatment + State.Id + Year, data = data)

summary(model2)
```

There is a significant decrease (beta = -4030.46, p < .001) in lung hospitalizations in treated states after the implementation of vaping laws in 2021.


# 5. Make regression table outputs

```{r}
model_results1 <- stargazer(model1, 
                            title="Regression Table 1", 
                            type = "html", 
                            dep.var.labels=c("Lung Hospitalizations"), 
                            covariate.labels=c("Treatment Applied (1)", 
                                               "Year", 
                                               "Treatment (1) x Year")) 

capture.output(model_results1, file = "~/Documents/GitHub/Research-Methods-II/HW 2/model1.html")
```


```{r}
model_results2 <- stargazer(model2, 
                            title="Regression Table 2", 
                            type = "html",
                            dep.var.labels=c("Lung Hospitalizations"),
                            covariate.labels=c("Post Treatment (1)",
                                               "Treatment Applied (1)",
                                               "State 2",
                                               "State 3",
                                               "State 4",
                                               "State 5",
                                               "State 6",
                                               "State 7",
                                               "State 8",
                                               "State 9",
                                               "State 10",
                                               "State 11",
                                               "State 12",
                                               "State 13",
                                               "State 14",
                                               "State 15",
                                               "State 16",
                                               "State 17",
                                               "State 18",
                                               "State 19",
                                               "State 20",
                                               "State 21",
                                               "State 22",
                                               "State 23",
                                               "State 24",
                                               "State 25",
                                               "State 26",
                                               "State 27",
                                               "State 28",
                                               "State 29",
                                               "State 30",
                                               "State 31",
                                               "State 32",
                                               "State 33",
                                               "State 34",
                                               "State 35",
                                               "State 36",
                                               "State 37",
                                               "State 38",
                                               "State 39",
                                               "State 40",
                                               "State 41",
                                               "State 42",
                                               "State 43",
                                               "State 44",
                                               "State 45",
                                               "State 46",
                                               "State 47",
                                               "State 48",
                                               "State 49",
                                               "State 50",
                                               "Year",
                                               "Post Treatment (1) x Treatment Applied (1)"
                                               )) 

capture.output(model_results2, file = "~/Documents/GitHub/Research-Methods-II/HW 2/model2.html")
```


# Can you reject the hypothesis that state fixed effects are all zero? 

```{r}
regTermTest(model2,
            "State.Id",
            method = c("LRT"),
            lrt.approximation = "satterthwaite") 
```

Yes!
